---
title: "Worksheet 0"
author: "Student Name (Student Number)"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Instructions

To complete this worksheet, add code as needed into the R code chunks given below. **Do not** delete the question text. All text should be in complete English sentences. Be sure to change the author of this file to reflect your name and student number. To properly see the questions, knit this .Rmd file to .pdf and view the output. 

This worksheet will not be submitted and does not count for marks.

\clearpage

# Exercises

***Make the following calculations in R: $2 + 2$, $4 - 1$, $3*4$, $3^5$, and $100/3$.***  

```{r}
#Delete this line (including the # symbol) and place your code here.
```

## Numeric Functions

R has functions for most common mathematical purposes. To use a function, we type the name of the function, followed by round brackets. The value in the brackets is called the *argument* of the function, and can be either a numeric value or a variable.

For example, we may calculate the square root of 15 by using the `sqrt` function.

***Calculate the square root of 15.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

Or, we can round a number to the nearest integer with the `round` function.

***Round 6.3 and 4.99 to their nearest integers.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

## Saving Variables

We may store variables in R as well. Once these variables are stored, we may read them again later, or do operations on them.

Variables are assigned by giving the name of the variable, followed by `<-`. Note that you may also use the `=` symbol, though the left-arrow is preferred.

***Save the value 5.73 as `x`. Then: print out the value of `x`. Finally, calculate `x + 1`, `x^2`, the square root of `x`, and round `x` to the nearest integer.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

## Vectors / Datasets

We may also create an entire dataset of values as an object called a *vector*. To do so, we list all of our values, separated by commas, and we wrap the entire list of values in the function `c()`.

For example, suppose that we have a class of 12 students. These students have the heights below, measured to the nearest cm:
$$167, \, 184, \, 175, \, 171, \, 168, \, 174, \, 163, \, 178, \, 158, \, 163, \, 165, \, 162$$

***Save the dataset above as a vector named `studentHeights`.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

***Print out the `studentHeights` vector.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

## Calculating a mean (average value)

R also contains a suite of functions for all common statistical calculations.

The first one we will look at is `mean`, which calculates the arithmetic mean (i.e., the "average") of a dataset.

***Calculate the mean of the `studentHeights` dataset.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

We can see that these students have a mean height of 169cm. This is much faster than doing the computations by hand, or with a calculator!

## Importing Data

In this course, all datasets will be provided to you in the form of a .csv (comma-separated value) file. This is the most common file type for dealing with datasets in statistics.

    
***Follow the data importing instructions provided in the RStudio Intro file to import the `Class` dataset, and copy the line of code that imports your dataset. This should be of the form ```Class <- read.csv(...)```. Do not include the View function.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

## Dataframes


***Print out the `Class` dataframe.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

Since this dataset is small, we can just print it out. However, many files will be too large to view in this way. In that case, you may try the `head` function, and the `str` function.

***Use the `head` function to print the first 6 rows of the `Class` dataframe.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

This is useful for checking how exactly the values in your dataset look.

The `str` function will show the *structure* of your dataframe: it will show all of the variables, along with their data type, as well as some example values.

***Use the `str` function to print the structure of the `Class` dataframe***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

We can see that the names are "character" data types, the heights are numeric (quantitative), and the majors are again "characters".

## The `$` operator

To refer to a particular variable in your dataframe, append a `$` to your dataframe, and then type the name of the variable.

***Use the `$` operator to print out the `Height` variable within the `Class` dataframe.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```


***Calculate the mean of the `Height` variable.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

## Creating a histogram

***Use the `hist` function to produce a simple histogram of the `Height` variable.***

```{r}
#Delete this line (including the # symbol) and place your code here.
```

We can see that the most common height for students in this class is around 170-180cm.




